---
title: "Graphs"
author: "Steven Lam & Alexander Lee"
date: "December 6, 2015"
output: word_document
---


Creates the charts for visualization of the data. Each chart shows the
frequency of each grade (A+, A, A-, etc.) and lines indicating the
average grade for that class. Classes are arranged by teacher.


```{r}
#re-import the data and whatnot so we can knit the .Rmd

library(readr)
setwd("../Data")

stat2010f <- read_csv("Grades2010f.csv")
stat2011f <- read_csv("Grades2011f.csv")
stat2012f <- read_csv("Grades2012f.csv")
stat2013f <- read_csv("Grades2013f.csv")
stat2014f <- read_csv("Grades2014f.csv")

stat2010s <- read_csv("Grades2010s.csv")
stat2011s <- read_csv("Grades2011s.csv")
stat2012s <- read_csv("Grades2012s.csv")
stat2013s <- read_csv("Grades2013s.csv")
stat2014s <- read_csv("Grades2014s.csv")
stat2015s <- read_csv("Grades2015s.csv")


#fall 2012
DeborahNolan2012f <- stat2012f[stat2012f$`Instructor Name` == "Deborah Nolan",]
FletcherIbser2012f <- stat2012f[stat2012f$`Instructor Name` == "Fletcher Ibser",]

#fall 2013
Ingileif2013f <- stat2013f[stat2013f$`Instructor Name` == "Ingileif Hallgrimsdottir",]
FletcherIbser2013f <- stat2013f[stat2013f$`Instructor Name` == "Fletcher Ibser",]

#fall 2014
BenoitDherin2014f <- stat2014f[stat2014f$`Instructor Name` == "Benoit Dherin",]
FletcherIbser2014f <- stat2014f[stat2014f$`Instructor Name` == "Fletcher Ibser",]

#spring 2012
DeborahNolan2012s <- stat2012s[stat2012s$`Instructor Name` == "Deborah Nolan",]
FletcherIbser2012s <- stat2012s[stat2012s$`Instructor Name` == "Fletcher Ibser",]

#spring 2014
BenoitDherin2014s <- stat2014s[stat2014s$`Instructor Name` == "Benoit Dherin",]
CariKaufman2014s <- stat2014s[stat2014s$`Instructor Name` == "Cari Kaufman",]

#spring 2015
DeborahNolan2015s <- stat2015s[stat2015s$`Instructor Name` == "Deborah Nolan",]
Ingileif2015s <- stat2015s[stat2015s$`Instructor Name` == "Ingileif Hallgrimsdottir",]


#GPA calculations
gpa <- c(12, 12, 11.1, 9.9, 9, 8.1, 6.9, 6, 5.1, 3.9, 3, 2.1, 0)
grade_names <- c("A+", "A", "A-", "B+", "B", "B-", 
                 "C+", "C", "C-", "D+", "D", "D-", "F")
names(gpa) <- grade_names
gpa <- gpa / 3


#loopdy loop - find the total grade distribution for all classes

total_sum <- double(13)
names(total_sum) <- grade_names
grades <- list(stat2010f, stat2011f, stat2012f, 
               stat2013f, stat2014f, stat2010s,  
               stat2011s, stat2012s, stat2013s, 
               stat2014s, stat2015s)

for(i in 1:length(grades)){
  
  #create a temporary variable for the data
  temp_data <- data.frame(grades[i])
  
  for(j in 1:13){
    
    total_sum[j] <- total_sum[j] + sum(temp_data[temp_data$Grade.Nm == grade_names[j], 
                                                 "Enrollment.Cnt"])
  }
  
}



#create a vector with the grade relative frequencies for sampling purposes
grade_samples <- rep(grade_names, times = total_sum)
```

#Data Visualization


##Cari Kaufman


####Cari Kaufman Fall 2010
```{r}
#stat2010f


barplot(height = stat2010f$`Enrollment Cnt`, names.arg = stat2010f$`Grade Nm`,
        main = "Cari Kaufman Fall 2010 Grade Distribution",
        xlab = "Grade Name", ylab = "Frequency", las = 1)

#create the GPA distribution of the class based on its
#`Enrollment Cnt`'s, used to find the mean/median
grade_samples <- rep(stat2010f$`Average Grade`, times = stat2010f$`Enrollment Cnt`)

#mean and median
average_grade <- mean(grade_samples)
median_grade <- median(grade_samples)

#round the mean and median
average_grade <- round(average_grade, 4)
median_grade <- round(median_grade, 4)


#create a legend that says what the lines mean
legend("topright", legend = c(paste0("mean = ", as.character(average_grade)), 
                              paste0("median = ", as.character(median_grade))))

```

####Cari Kaufman Fall 2011
```{r}
#stat2011f

barplot(height = stat2011f$`Enrollment Cnt`, names.arg = stat2011f$`Grade Nm`,
        main = "Cari Kaufman Fall 2011 Grade Distribution",
        xlab = "Grade Name", ylab = "Frequency", las = 1)

#create the GPA distribution of the class based on its
#`Enrollment Cnt`'s, used to find the mean/median
grade_samples <- rep(stat2011f$`Average Grade`, times = stat2011f$`Enrollment Cnt`)

#mean and median
average_grade <- mean(grade_samples)
median_grade <- median(grade_samples)

#round the mean and median
average_grade <- round(average_grade, 4)
median_grade <- round(median_grade, 4)


#create a legend that says what the lines mean
legend("topright", legend = c(paste0("mean = ", as.character(average_grade)), 
                              paste0("median = ", as.character(median_grade))))
```

####Cari Kaufman Spring 2014
```{r}
#stat2014s

barplot(height = stat2014s$`Enrollment Cnt`, names.arg = stat2014s$`Grade Nm`,
        main = "Cari Kaufman Spring 2014 Grade Distribution",
        xlab = "Grade Name", ylab = "Frequency", las = 1)

#create the GPA distribution of the class based on its
#`Enrollment Cnt`'s, used to find the mean/median
grade_samples <- rep(stat2014s$`Average Grade`, times = stat2014s$`Enrollment Cnt`)

#mean and median
average_grade <- mean(grade_samples)
median_grade <- median(grade_samples)

#round the mean and median
average_grade <- round(average_grade, 4)
median_grade <- round(median_grade, 4)


#create a legend that says what the lines mean
legend("topright", legend = c(paste0("mean = ", as.character(average_grade)), 
                              paste0("median = ", as.character(median_grade))))
```

####Cari Kaufman - Totals for All Years
```{r}
#stat2010f
#stat2011f
#CariKaufman2014s


#combine the grade data for all of
#Cari Kaufman's classes

grades <- list(DeborahNolan2012f, DeborahNolan2012s, DeborahNolan2015s)

total_sum <- double(13)
names(total_sum) <- grade_names

for(i in 1:length(grades)){
  
  #create a temporary variable for the data
  temp_data <- data.frame(grades[i])
  
  for(j in 1:13){
    
    total_sum[j] <- total_sum[j] + sum(temp_data[temp_data$Grade.Nm == grade_names[j],
                                                 "Enrollment.Cnt"])
    
  }
  
}

total_sum <- total_sum[total_sum != 0]
gpa_values <- unname(gpa[intersect(names(gpa), names(total_sum))])


barplot(height = total_sum, names.arg = names(total_sum),
        main = "Cari Kaufman Grade Distribution for All Classes",
        xlab = "Grade Name", ylab = "Frequency", las = 1)

#create the GPA distribution of the class based on its
#`Enrollment Cnt`'s, used to find the mean/median
grade_samples <- rep(gpa_values, times = total_sum)

#mean and median
average_grade <- mean(grade_samples)
median_grade <- median(grade_samples)

#round the mean and median
average_grade <- round(average_grade, 4)
median_grade <- round(median_grade, 4)


#create a legend that says what the lines mean
legend("topright", legend = c(paste0("mean = ", as.character(average_grade)), 
                              paste0("median = ", as.character(median_grade))))

```


```{r}
#2013s

barplot(height = stat2013s$`Enrollment Cnt`, names.arg = stat2013s$`Grade Nm`,
        main = "Haiyan Huang Spring 2013 Grade Distribution",
        xlab = "Grade Name", ylab = "Frequency", las = 1)

#create the GPA distribution of the class based on its
#`Enrollment Cnt`'s, used to find the mean/median
grade_samples <- rep(stat2013s$`Average Grade`, times = stat2013s$`Enrollment Cnt`)

#mean and median
average_grade <- mean(grade_samples)
median_grade <- median(grade_samples)

#round the mean and median
average_grade <- round(average_grade, 4)
median_grade <- round(median_grade, 4)


#create a legend that says what the lines mean
legend("topright", legend = c(paste0("mean = ", as.character(average_grade)), 
                              paste0("median = ", as.character(median_grade))))
```

####Deborah Nolan - All 
```{r}
grades <- list(DeborahNolan2012f, DeborahNolan2012s, DeborahNolan2015s)

total_sum <- double(13)
names(total_sum) <- grade_names

for(i in 1:length(grades)){
  
  #create a temporary variable for the data
  temp_data <- data.frame(grades[i])
  
  for(j in 1:13){
    
    total_sum[j] <- total_sum[j] + sum(temp_data[temp_data$Grade.Nm == grade_names[j],
                                                 "Enrollment.Cnt"])
    
  }
  
}

total_sum <- total_sum[total_sum != 0]
gpa_values <- unname(gpa[intersect(names(gpa), names(total_sum))])


barplot(height = total_sum, names.arg = names(total_sum),
        main = "Deborah Nolan Grade Distribution for All Classes",
        xlab = "Grade Name", ylab = "Frequency", las = 1)

#create the GPA distribution of the class based on its
#`Enrollment Cnt`'s, used to find the mean/median
grade_samples <- rep(gpa_values, times = total_sum)

#mean and median
average_grade <- mean(grade_samples)
median_grade <- median(grade_samples)

#round the mean and median
average_grade <- round(average_grade, 4)
median_grade <- round(median_grade, 4)


#create a legend that says what the lines mean
legend("topright", legend = c(paste0("mean = ", as.character(average_grade)), 
                              paste0("median = ", as.character(median_grade))))
```

####Fletcher Ibser - All
```{r}
grades <- list(FletcherIbser2012f, FletcherIbser2012s, 
               FletcherIbser2013f, FletcherIbser2014f)

total_sum <- double(13)
names(total_sum) <- grade_names

for(i in 1:length(grades)){
  
  #create a temporary variable for the data
  temp_data <- data.frame(grades[i])
  
  for(j in 1:13){
    
    total_sum[j] <- total_sum[j] + sum(temp_data[temp_data$Grade.Nm == grade_names[j],
                                                 "Enrollment.Cnt"])
    
  }
  
}

total_sum <- total_sum[total_sum != 0]
gpa_values <- unname(gpa[intersect(names(gpa), names(total_sum))])


barplot(height = total_sum, names.arg = names(total_sum),
        main = "Fletcher Ibser Grade Distribution for All Classes",
        xlab = "Grade Name", ylab = "Frequency", las = 1)

#create the GPA distribution of the class based on its
#`Enrollment Cnt`'s, used to find the mean/median
grade_samples <- rep(gpa_values, times = total_sum)

#mean and median
average_grade <- mean(grade_samples)
median_grade <- median(grade_samples)

#round the mean and median
average_grade <- round(average_grade, 4)
median_grade <- round(median_grade, 4)


#create a legend that says what the lines mean
legend("topright", legend = c(paste0("mean = ", as.character(average_grade)), 
                              paste0("median = ", as.character(median_grade))))
```

####Benoit Dherin - All
```{r}
grades <- list(stat2013f, BenoitDherin2014s, BenoitDherin2014f)

total_sum <- double(13)
names(total_sum) <- grade_names

for(i in 1:length(grades)){
  
  #create a temporary variable for the data
  temp_data <- data.frame(grades[i])
  
  for(j in 1:13){
    
    total_sum[j] <- total_sum[j] + sum(temp_data[temp_data$Grade.Nm == grade_names[j],
                                                 "Enrollment.Cnt"])
    
  }
  
}

total_sum <- total_sum[total_sum != 0]
gpa_values <- unname(gpa[intersect(names(gpa), names(total_sum))])


barplot(height = total_sum, names.arg = names(total_sum),
        main = "Benoit Dherin Grade Distribution for All Classes",
        xlab = "Grade Name", ylab = "Frequency", las = 1)

#create the GPA distribution of the class based on its
#`Enrollment Cnt`'s, used to find the mean/median
grade_samples <- rep(gpa_values, times = total_sum)

#mean and median
average_grade <- mean(grade_samples)
median_grade <- median(grade_samples)

#round the mean and median
average_grade <- round(average_grade, 4)
median_grade <- round(median_grade, 4)


#create a legend that says what the lines mean
legend("topright", legend = c(paste0("mean = ", as.character(average_grade)), 
                              paste0("median = ", as.character(median_grade))))
```

####Ingileif Hallgrimsdottir - All
```{r}
grades <- list(Ingileif2013f, Ingileif2015s)

total_sum <- double(13)
names(total_sum) <- grade_names

for(i in 1:length(grades)){
  
  #create a temporary variable for the data
  temp_data <- data.frame(grades[i])
  
  for(j in 1:13){
    
    total_sum[j] <- total_sum[j] + sum(temp_data[temp_data$Grade.Nm == grade_names[j],
                                                 "Enrollment.Cnt"])
    
  }
  
}

total_sum <- total_sum[total_sum != 0]
gpa_values <- unname(gpa[intersect(names(gpa), names(total_sum))])


barplot(height = total_sum, names.arg = names(total_sum),
        main = "Ingileif Hallgrimsdottir Grade Distribution for All Classes",
        xlab = "Grade Name", ylab = "Frequency", las = 1)

#create the GPA distribution of the class based on its
#`Enrollment Cnt`'s, used to find the mean/median
grade_samples <- rep(gpa_values, times = total_sum)

#mean and median
average_grade <- mean(grade_samples)
median_grade <- median(grade_samples)

#round the mean and median
average_grade <- round(average_grade, 4)
median_grade <- round(median_grade, 4)


#create a legend that says what the lines mean
legend("topright", legend = c(paste0("mean = ", as.character(average_grade)), 
                              paste0("median = ", as.character(median_grade))))
```

####Philip Spector - All
```{r}
grades <- list(stat2010s, stat2011s) 

total_sum <- double(13)
names(total_sum) <- grade_names

for(i in 1:length(grades)){
  
  #create a temporary variable for the data
  temp_data <- data.frame(grades[i])
  
  for(j in 1:13){
    
    total_sum[j] <- total_sum[j] + sum(temp_data[temp_data$Grade.Nm == grade_names[j],
                                                 "Enrollment.Cnt"])
    
  }
  
}

total_sum <- total_sum[total_sum != 0]
gpa_values <- unname(gpa[intersect(names(gpa), names(total_sum))])


barplot(height = total_sum, names.arg = names(total_sum),
        main = "Philip Spector Grade Distribution for All Classes",
        xlab = "Grade Name", ylab = "Frequency", las = 1)

#create the GPA distribution of the class based on its
#`Enrollment Cnt`'s, used to find the mean/median
grade_samples <- rep(gpa_values, times = total_sum)

#mean and median
average_grade <- mean(grade_samples)
median_grade <- median(grade_samples)

#round the mean and median
average_grade <- round(average_grade, 4)
median_grade <- round(median_grade, 4)


#create a legend that says what the lines mean
legend("topright", legend = c(paste0("mean = ", as.character(average_grade)), 
                              paste0("median = ", as.character(median_grade))))
```

####All Classes
```{r}
grades <- list(stat2010f, stat2011f, stat2012f, 
               stat2013f, stat2014f, stat2010s,  
               stat2011s, stat2012s, stat2013s, 
               stat2014s, stat2015s)

total_sum <- double(13)
names(total_sum) <- grade_names

for(i in 1:length(grades)){
  
  #create a temporary variable for the data
  temp_data <- data.frame(grades[i])
  
  for(j in 1:13){
    
    total_sum[j] <- total_sum[j] + sum(temp_data[temp_data$Grade.Nm == grade_names[j],
                                                 "Enrollment.Cnt"])
    
  }
  
}

total_sum <- total_sum[total_sum != 0]
gpa_values <- unname(gpa[intersect(names(gpa), names(total_sum))])


barplot(height = total_sum, names.arg = names(total_sum),
        main = "Grade Distribution for All Classes for All Years",
        xlab = "Grade Name", ylab = "Frequency", las = 1)

#create the GPA distribution of the class based on its
#`Enrollment Cnt`'s, used to find the mean/median
grade_samples <- rep(gpa_values, times = total_sum)

#mean and median
average_grade <- mean(grade_samples)
median_grade <- median(grade_samples)

#round the mean and median
average_grade <- round(average_grade, 4)
median_grade <- round(median_grade, 4)


#create a legend that says what the lines mean
legend("topright", legend = c(paste0("mean = ", as.character(average_grade)), 
                              paste0("median = ", as.character(median_grade))))

total_prop <- prop.table(total_sum)
sum(head(total_prop, 3))
sum(total_prop[4:6])
```